---
title: "Penguin quarto"
format: html
editor: visual
---

# initialisation

```{r}
#|echo: false
library(tidyverse)
library(here)
library(palmerpenguins)
library(janitor)
```

## renv

allows us to keep track of installed libraries and their versions

ensures that other people can run everything correctly

```{r}
renv::init()
```

```{r}
renv::snapshot()
```

You can add options to executable code like this

```{r}
colnames(penguins_raw)
```

```{r}
write.csv(penguins_raw, here("data","penguins_raw.csv"))
```

here is a relative path, goes to the current folder. Allows it to be used on other

# modifying columns

```{r}
penguins_clean <- penguins_raw %>% 
  select(-Comments) %>% 
  select(-starts_with("Delta")) %>% 
  clean_names()

colnames(penguins_clean)

write.csv(penguins_clean, here("data","penguins_clean.csv"))
```

## as a function

shorten_species()

```{r}
cleaning_columns <- function(raw_data){
  raw_data %>% 
    select(-Comments) %>% 
    select(-starts_with("Delta")) %>% 
    clean_names()
}

baby_boi <- cleaning_columns(penguins_raw)
```

## cleaning up NAs

in this we are creating a new variable which has the NA removed, if we remove them from all comuns at once then it will remove all observations with an NA in any column

```{r}
# only removes data from species and flipper length
penguins_flippers <- penguins_clean %>%
  select(species, flipper_length_mm) %>%
  drop_na() %>% 
  shorten_species()
```

# reproducable figures

## exploratory figures

shows the raw data, always have to be shown when working with data

classic graphs and stuff like that

nothing fitted to or hypothesis testy

### boxplot

```{r}
species_colours <- c("Adelie" = "darkorange",
                       "Chinstrap" = "purple",
                       "Gentoo" = "cyan4")

BoxFlip <- ggplot(data = penguins_flippers, 
                  aes(x = species, 
                      y = flipper_length_mm)) +
  geom_boxplot(aes(colour = species)
               ,show.legend = FALSE,
               width = 0,5) +
  geom_jitter(aes(colour = species), 
              position = position_jitter(seed = 0)) +
  theme_bw()+
  scale_colour_manual(values = species_colours) +
  labs(x = "species", y = "Flipper Length (mm)")

BoxFlip
```
